# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
Input data is a ZIP file containing a single CSV file.
First uncompress that file into a temporary directory "tmp"

```{r unzip, echo=TRUE}
unzip(zipfile="activity.zip", overwrite=TRUE, exdir="tmp")
```

Then read the data into the *inputdata* variable

```{r inputdata, echo=TRUE}
inputdata <- read.csv("tmp/activity.csv", head = TRUE)
```

## What is mean total number of steps taken per day?
First number of steps by day is calculated:

```{r stepsbyday, echo=TRUE}
stepsbyday <- aggregate( steps~date, inputdata, sum)
```

```{r histstepsbyday, echo=FALSE}
hist(stepsbyday$steps, main = "Histogram of total steps by day", xlab="Steps by day")
```

Then mean and median of the step array:

```{r meanstepsbyday, echo=TRUE}
mean(stepsbyday$steps)
median(stepsbyday$steps)
```

## What is the average daily activity pattern?
```{r meandailyactivity, echo=TRUE}
meandailyactivity <- aggregate( steps~interval, inputdata, mean)

plot(x=meandailyactivity, type='l')
```


## Imputing missing values
1. Total number of missing values in the dataset
```{r total missing values, echo=TRUE}
numberofmissingvalues <- nrow(inputdata[is.na(inputdata$steps),])
numberofmissingvalues
```

2. Strategy for filled-in missing number of steps.
Any row with NA number of steps will be replaced with the mean of
the corresponding 5-minute interval.

3. New dataset based on (2)

```{r fill-in missing values, echo=TRUE}
for (i in seq(nrow(inputdata))) {
    if (is.na(inputdata[i,"steps"])) {
        interval <- inputdata[i,"interval"]
        inputdata[i,"steps"] <-
          meandailyactivity[meandailyactivity$interval == interval, "steps"]
    }
}
```
4. Histogram of total number of steps each day, then mean and median for this

```{r newstepsbyday, echo=TRUE}
stepsbyday <- aggregate( steps~date, inputdata, sum)
```

```{r newhiststepsbyday, echo=FALSE}
hist(stepsbyday$steps, main = "Histogram of total steps by day", xlab="Steps by day")
```

Then mean and median of the new step array.
Note that mean shall not be affected but median might have changed
with this strategy:

```{r newmeanstepsbyday, echo=TRUE}
mean(stepsbyday$steps)
median(stepsbyday$steps)



## Are there differences in activity patterns between weekdays and weekends?
